Modelling low Mach number stellar hydrodynamics with MAESTROeX

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
added citation for massive star rotation rates

The massive stars that we are interested in modelling are believed to be relatively slow rotators\MarginPar{insert some citation/quantify this}, so it is likely that one of the simpler approaches outlined above should be sufficient. However, for modelling e.g.~type I X-ray bursts, it may be that a more sophisticated approach is needed. Type I X-ray bursts are produced by thermonuclear deflagrations in the liquid surface layers of fast rotating neutron stars. These stars typically rotate at frequencies of $\sim$300\textendash600 Hz, a significant fraction of the stars' breakup velocity. This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meridional angle. To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meridional angle $p_0(r)\rightarrow p_0(r, \theta)$.

>>>>

The massive stars that we are interested in modelling are believed to be relatively slow rotators, with the rotation rate in the core prior to collapse no more than a few percent of the Kepler frequency \cite{heger2004presupernova}, so it is likely that one of the simpler approaches outlined above should be sufficient. However, for modelling e.g.~type I X-ray bursts, it may be that a more sophisticated approach is needed. Type I X-ray bursts are produced by thermonuclear deflagrations in the liquid surface layers of fast rotating neutron stars. These stars typically rotate at frequencies of $\sim$300\textendash600 Hz, a significant fraction of the stars' breakup velocity. This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meridional angle. To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meridional angle $p_0(r)\rightarrow p_0(r, \theta)$.
